\documentstyle{Rosetta}
\begin{document}
   \RosTopic{General}
   \RosTitle{Retrograde Vocabulary Van Dale E~--~N}
   \RosAuthor{Jeroen Medema, Harm Smit}
   \RosDocNr{193}
   \RosDate{\today}
   \RosStatus{Informal}
   \RosSupersedes{-}
   \RosDistribution{Harm Smit, Jeroen Medema}
   \RosClearance{Project}
   \RosKeywords{Dictionary, Retrograde vocabulary, E~--N}
   \MakeRosTitle
%
%
\setlength{\parindent}{0mm}
\setlength{\parskip}{2mm}

\section{Introduction}
 The following pages contain the vocabulary of the E~--~N (= English -- Dutch) 
 dictionary of 
 Van Dale (\cite{ma:groot}). The ordering of the words is retrograde (which 
 means that they are not ordered with respect to the first characters of a 
 word -as in normal dictionaries- but with respect to the last ones). 
 This list contains all the main entries as given in the 
 N~--~N. It consist of exactly 86281 entries (count it yourself).

 It was made to be used by linguists and dictionary programmers when a set 
 of words with a special ending is needed (especially for the morphological 
 part). 
\section{Vocabulary order}
 The following constraints apply to the ordering:
 \begin{enumerate}
  \item characters with an accent (as e.g. \^{e}, \~{n}, \aa, \"{o}) are 
  considered to be equal to the same character without the accent.
  \item words containing the string ``(-)'' are considered to be equal to the 
  same word without that string.
  \item words containing {\em other} characters than letters, ``-'', ``~'', and 
  ``.'' are considered to be equal to the same word without those characters.
  \item the ordering is done by smallest first; that is: first ``~'', then 
  ``-'', next ``.'', followed by the letters.
  \item the upper and lower case form of a letter are considered to be equal.
 \end {enumerate}
 The words are split up in twenty-eight parts: for every letter one part and 
 two additional for the ``-'' and ``.'' (no word ends with a space). Each part 
 starts at a new page.
\section{Realisation}
 Here, I refer to \cite{med:crv}.

\begin{thebibliography}{AAA99a}
 \bibitem[MAR84]{ma:groot} W. Martin, ed. {\em Groot woordenboek 
  Engels -- Nederlands}, Van Dale Lexicografie, 1984.
 \bibitem[MED87]{med:crv} Jeroen Medema, {\em System to Create Retrograde 
  Dictionaries}, {\bf Rosetta report 190}, Philips Research labs, 1987.
\end{thebibliography}
\end{document}
